---
title: "L05 Toolbox"
subtitle: "Data Visualization (STAT 302)"
author: "Martha Eichlersmith"
date: 2020-04-19
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: "hide"
---

## Overview

The goal of this lab is to explore more useful plots in `ggplot2`. Specifically we will be focusing on surface plots and geospatial plots (maps).

Challenges are not mandatory for students to complete. We highly recommend students attempt them though. *We would expect graduate students to attempt the challenges.*

```{r L05-setup, message=FALSE, warning=FALSE}
# Load package(s)

library(tidyverse)
library(gridExtra)

 #for geom_hex 
library(hexbin)
library(maps)

#https://github.com/hrbrmstr/statebins
library(statebins)

#color pallet https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html
library(viridis) 

#for geom_sf; will not work with discrete fill unless the package is loaded 
library(sf)

knitr::opts_chunk$set(dpi = 300) 
```

## Datasets

We'll be using data from the `blue_jays.rda`, `cdc.txt`, and the `US_income.rda` datasets which are already in the `/data` subdirectory in our **data_vis_labs** project. 

```{r LO4-load-data}
# Load datasets
load(file = "data/blue_jays.rda")
load(file = "data/US_income.rda")

# Read in the cdc dataset
cdc <- read_delim(
  file = "data/cdc.txt"
  , delim = "|"
  #add col types 
  , col_types = cols(
                      genhlth = col_character(),
                      exerany = col_double(),
                      hlthplan = col_double(),
                      smoke100 = col_double(),
                      height = col_double(),
                      weight = col_double(),
                      wtdesire = col_double(),
                      age = col_double(),
                      gender = col_character()
                    )
  ) %>%
  mutate(genhlth = factor(genhlth,
    levels = c("excellent", "very good", "good", "fair", "poor")
  ))
```


## Exercises

Complete the following exercises.

<br>

### Exercise 1

The following plot uses the `blue_jays.rda` dataset.


```{r LO5-EX1}
#body mass by head
ggplot(blue_jays, aes(Mass, Head))+
  #scatter plot 
  geom_point(alpha = 1/3, size = 1.5)+
  #2D density plot 
  geom_density_2d(
    #set binwidth for density plot
      binwidth = 0.004
    , color = "black"
    , size = 0.4
    )+
  #set limits on x-axis 
  xlim(57, 82)+
  # add labels 
  labs(
      x = "Body Mass (g)"
    , y = "Head length (mm)"
  )+
  #add minimal theme; white background on plot
  theme_minimal()
```


<br>

### Exercise 2

The following plots uses the `cdc` dataset

#### Plot 1


```{r LO5-EX2-plot1}
cdc_plot <- ggplot(cdc, aes(height, weight))+
  #set axis titles 
  labs(
      x = "Height (in)"
    , y = "Weight (lbs)"
  )+
  #add minimal theme; white background on plot
  theme_minimal()

#2d density plot with hexagons
cdc_plot + geom_hex(bins=35)
```



<br>

#### Plot 2


```{r LO5-EX2-plot2}
cdc_plot+
  stat_density_2d(
    #make shading based on density 
      aes(fill=after_stat(level))
    #fill density plot (solid not just lines)
    , geom="polygon"
    #don't show legend 
    , show.legend = FALSE
    )+
  #separate plots for gender
  facet_wrap(~gender)
```



<br>

### Exercise 3

the following plot is made using the `maps` library.  

```{r LO5-EX3}
#the best state in the union 
mn <- map_data("county", "minnesota") %>% 
  select(long, lat, group, id=subregion)

ggplot(mn, aes(x=long, y=lat))+
  #add title+
  ggtitle("Minnesota")+
  #draws the shapes 
  geom_polygon(aes(group=group), fill="white", color="grey35")+
  #helps size correctely 
  coord_quickmap()+
  #shows just the plotted features 
  theme_void()
```

<br>  

### Exercise 4

Using `US_income` dataset, recreate the following graphics as precisely as possible.

<br>


```{r LO5-EX4-data}
# Setting income levels
US_income <- mutate(
  US_income,
  income_bins = cut(
    ifelse(is.na(median_income), 25000, median_income),
    breaks = c(0, 40000, 50000, 60000, 70000, 80000),
    labels = c("< $40k", "$40k to $50k", "$50k to $60k", "$60k to $70k", "> $70k"),
    right = FALSE
  )
)

```


<br>


#### Plot 1


```{r LO5-EX4-plot1}
ggplot(US_income)+
  # 'simple features', need sf package
  geom_sf(
      aes(
        #specify geometry shape to be used 
          geometry = geometry
        #fill based on median income 
        , fill = income_bins
        )
    #set boundary color
    , color = "grey80"
    #set boundary line size
    , size=0.2
    )+
  #scale color fill to viridis color pallet 
  viridis::scale_fill_viridis(
    #change to continous scale
      discrete = TRUE
    #change legend name
    , name="Median\nIncome"
    )+
  #helps size correctely 
  coord_sf()+
  #shows just the plotted features 
  theme_void()
```



<br>

#### Plot 2

```{r LO5-EX4-plot2}
ggplot(US_income, aes(state=name, fill=income_bins))+
  geom_statebins()+
  #make each bin an equal square
#  coord_equal()+
  #scale color fill to viridis color pallet 
  viridis::scale_fill_viridis(
    #change to continous scale
      discrete = TRUE
    #change legend name
    , name="Median\nIncome"
    )+
  theme_statebins()
```



<br>

## Challenge(s)

Using the `tidycensus` package and few others, try to create a map like below using [these directions](https://walkerke.github.io/tidycensus/articles/spatial-data.html). Try using a different geographical area and a different variable from the ACS.

<br>

