---
title: "L01 Introduction"
subtitle: "Data Visualization (STAT 302)" 
date: 2020-04-12
author: "Martha Eichlersmith"
output:
  html_document:
    toc: true
    toc_float: true
    highlight: "tango"
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

## Overview

The goals of this lab are to (1) ensure that the major software for this course is properly installed and functional, (2) develop and follow a proper workflow, and (3) work together to construct a few plots to explore a dataset using `ggplot2` --- demonstration of the utility and power of `ggplot2`. 

**Don't worry if you cannot do everything here by yourself.** You are just getting started and the learning curve is steep, but remember that the instructional team and your classmates will be there to provide support. Persevere and put forth an honest effort and this course will payoff.

**Load Packages** `tidyverse`, `ggstance`, `skimr` 

```{r load-packages}
# Load package(s) 
suppressPackageStartupMessages(library(tidyverse))
#suppressPackageStartupMessages(library(ggstance))
suppressPackageStartupMessages(library(skimr))
```

<br>

## Dataset

We'll be using data from the [**lego** package](https://github.com/seankross/lego) which is already in the `/data` subdirectory, along with many other processed datasets, as part of the zipped folder for this lab. 
<br> 

## Exercise 1

Let's look at some interesting patterns in the history of LEGO! We'll be using data from the [**lego** package](https://github.com/seankross/lego) located `data/legosets.rda`. We will work through this exercise together in class.

```{r loaddata}
load(file = "data/legosets.rda")
```


### 1a Inspect the data

The **lego** package provides a helpful dataset some interesting variables. Let's take a quick look at the data.

```{r dataset}
#print the top of the data sets
#head(legosets)

#focus of the variables 
glimpse(legosets)

#short eda
skim_without_charts(legosets)
```
  
Notice there are a lot of missing variables, especially when it comes to pricing - this will be important for when we calculate the means.  

<br>

### 1b Price per year

First, let's look at the average cost of LEGO sets over time. The main variable of interest here is `USD_MSRP`, or the manufacturer's suggested retail price in constant dollars (i.e. not adjusted for inflation).  

```{r priceVtime}
#`drop` NA's at top  
df.1b <- legosets %>% 
  drop_na(USD_MSRP) %>% 
  group_by(Year) %>% 
  summarize(
    AveragePrice = mean(USD_MSRP)
  )

#`remove` NA's in mean function
#equivalent to `drop` NAs at top for this task 
#df.1b <- legosets %>%
#  group_by(Year) %>%
#  summarize(
#    AverageCost = mean(USD_MSRP, na.rm=TRUE)
#    )

ggplot(df.1b, aes(Year, AveragePrice))+
  geom_point(size=1.25)+
  geom_line(size=1)+
  ggtitle("LEGO Sets: Average Price vs Year")+
  theme_minimal()+
  labs( #lables on the plot
     x = "Year"
   , y = "Average Price (USD)"
   , subtitle = "sets without price are removed from calculation of average price"
   , caption = "Source: LEGO"
    )
```

Is the increase in price simply due to inflation?  Could get data from federal resersve and plot inflation-adjusted price vs year.  
<br><br>



### 1c Pieces per year

Next, let’s look at how the number of pieces per set has changed over time. Because Duplo sets are much smaller (since they’re designed for toddlers), we’ll make a special indicator variable for them.


```{r data-wrangling, echo=FALSE, eval=FALSE}
#not used but works 
#subset large df (just to make it simpler)
#don't want to amend actual legosets df 
legosubset1c<- select(legosets, Year, Theme, Pieces)

#function: to make theme into binary "Duplo" and "Not Duplo" variable 
myefunc_IsDuplo <- function(input){
  if (input == "Duplo" ){output= "Duplo"}
  else {output = "Not Duplo"}
  return(output)
}

#add binary duplo variable to df
legosubset1c$IsDuplo <- sapply(legosubset1c$Theme, myefunc_IsDuplo)

#create df with avg pieces grouped by year and duplo status
df.1c <- legosubset1c  %>%
  group_by(Year, IsDuplo) %>%
  summarize(
    AveragePieces = #name the column
      mean( #find the mena
        Pieces #what you wnat to find the mean of 
        , na.rm=TRUE #ignore (remove) NA's
        )
    )
```

#### All Years 
```{r piecesVtime}

df.1c <- legosets %>% 
  mutate(IsDuplo = ifelse(Theme == "Duplo", "Duplo", "Lego")) %>%
  drop_na(Pieces) %>% 
  group_by(Year, IsDuplo) %>% 
  summarize(
     AveragePieces = mean(Pieces)
    ,num_sets = n()
    )

#same plot 
ggplot(df.1c, aes(Year, AveragePieces, color=IsDuplo))+
  geom_point()+
  geom_line()+
  ggtitle("Lego Sets: Average Pieces vs Year")+
  labs( #lables on the plot
     x = "Year"
   , y = "Average Pieces"
   , subtitle = "sets without pieces are removed from calculation of average pieces"
   , caption = "Source: LEGO"
    )
  
#facet plots
ggplot(df.1c, aes(Year, AveragePieces, color=IsDuplo))+
  facet_wrap(~IsDuplo, ncol=2, scales="fixed")+
  geom_point()+
  geom_line()+
  ggtitle("Lego Sets: Average Pieces vs Year")+
  labs( #lables on the plot
     x = "Year"
   , y = "Average Pieces"
   , subtitle = "sets without pieces are removed from calculation of average pieces"
   , caption = "Source: LEGO"
    )+
  theme(
    legend.position = "none"
  )
```
<br> 

#### Data after 1985
```{r piecesVtime-after1985}

df.1c <- legosets %>% 
  mutate(IsDuplo = ifelse(Theme == "Duplo", "Duplo", "Lego")) %>%
  drop_na(Pieces) %>% 
  filter(Year > 1985) %>%
  group_by(Year, IsDuplo) %>% 
  summarize(
     AveragePieces = mean(Pieces)
    ,num_sets = n()
    )

#same plot 
ggplot(df.1c, aes(Year, AveragePieces, color=IsDuplo))+
  geom_point()+
  geom_line()+
  ggtitle("Lego Sets: Average Pieces vs Year")+
  labs( #lables on the plot
     x = "Year"
   , y = "Average Pieces"
   , subtitle = "sets without pieces are removed from calculation of average pieces"
   , caption = "Source: LEGO"
    )
  
#facet plots
ggplot(df.1c, aes(Year, AveragePieces, color=IsDuplo))+
  facet_wrap(~IsDuplo, ncol=2, scales="fixed")+
  geom_point()+
  geom_line()+
  ggtitle("Lego Sets: Average Pieces vs Year")+
  labs( #lables on the plot
     x = "Year"
   , y = "Average Pieces"
   , subtitle = "sets without pieces are removed from calculation of average pieces"
   , caption = "Source: LEGO"
    )+
  theme(
    legend.position = "none"
  )
```

<br> 

### 1d LEGO set themes

In the 1990s, [LEGO began partnering with famous brands and franchises](https://www.cnbc.com/2018/04/27/lego-marketing-strategy-made-it-world-favorite-toy-brand.html) to boost its own sales. First, let's see how many different "themes" LEGO now offers:

```{r numberofthemes}
length(unique(legosets$Theme))

legosets %>% 
  distinct(Theme) %>% 
  summarize ( n_themes = n())
```


```{r themes}
#counts of the themes 
df.1d <- legosets %>% 
  #count data, automatically labeled 'n' 
  count(Theme, sort = TRUE
        , name ="theme_count" #personalize name if wnat 
        )  %>% 
  #turning character chr into factor fct -- meaning a categorical variable 
  mutate(Theme = fct_inorder(Theme, ordered = TRUE)) 



df.1d %>% 
  filter(theme_count > 150, Theme != 'Duplo') %>% #only look at top ones
  ggplot(aes(x=fct_rev(Theme), y= theme_count)) +
    geom_col()+
    coord_flip()+
    labs(y="Number of Sets", x=NULL)


```

