---
title: "L03 ggplot II"
subtitle: "Data Visualization (STAT 302)"
author: "Martha Eichlersmith"
date: 2020-04-19
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: "hide"
---

```{r LO3-load-packages, warning=FALSE, message=FALSE}
# Load package(s)
library(tidyverse)
library(lubridate) #for ymd() function
```
## Overview

The goal of this lab is to continue the process of unlocking the power of `ggplot2` through constructing and experimenting with a few basic plots.

## Datasets

We'll be using data from the `BA_degrees.rda` and `dow_jones_industrial.rda` datasets which are already in the `/data` subdirectory in our **data_vis_labs** project. Below is a description of the variables contained in each dataset.

`BA_degrees.rda`

* `field` - field of study
* `year_str` - academic year (e.g. 1970-71)
* `year` - closing year of academic year
* `count` - number of degrees conferred within a field for the year
* `perc` - field's percentage of degrees conferred for the year

```{r LO3-load-date-ba-degrees}
load(file="data/BA_degrees.rda")
```


`dow_jones_industrial.rda`

* `date` - date
* `open` - Dow Jones Industrial Average at open
* `high` - Day's high for the Dow Jones Industrial Average
* `low` - Day's low for the Dow Jones Industrial Average
* `close` - Dow Jones Industrial Average at close
* `volume` - number of trades for the day

```{r LO3-load-date-djia}
load(file="data/dow_jones_industrial.rda")
```


We'll also be using a subset of the BRFSS (Behavioral Risk Factor Surveillance System) survey collected annually by the Centers for Disease Control and Prevention (CDC). The data can be found in the provided `cdc.txt` file --- place this file in your `/data` subdirectory. The dataset contains 20,000 complete observations/records of 9 variables/fields, described below.

* `genhlth` - How would you rate your general health? (excellent, very good, good, fair, poor)
* `exerany` - Have you exercised in the past month? (`1 = yes`, `0 = no`) 
* `hlthplan` - Do you have some form of health coverage? (`1 = yes`, `0 = no`) 
* `smoke100` - Have you smoked at least 100 cigarettes in your life time? (`1 = yes`, `0 = no`) 
* `height` -  height in inches
* `weight` - weight in pounds
* `wtdesire` - weight desired in pounds
* `age` - in years
* `gender` - `m` for males and `f` for females


```{r LO3-load-data, message=FALSE}
cdc <- read_delim(file = "data/cdc.txt", delim = "|") %>%
  mutate(genhlth = factor(genhlth,
    levels = c("excellent", "very good", "good", "fair", "poor")
  ))
```



## Exercises


### Exercise 1

The following exercises use the `BA_degrees` data set. 


```{r LO3-EX1-data}
# Wrangling for plotting
ba_dat <- BA_degrees %>% 
  # mean % per field
  group_by(field) %>% 
  mutate(mean_perc = mean(perc)) %>% 
  # Only fields with mean >= 5%
  filter(mean_perc >= 0.05) %>%
  # Organizing for plotting
  arrange(desc(mean_perc), year) %>% 
  ungroup() %>%
  mutate(field = fct_inorder(field))

#take a look at the data to see variable names and types
head(ba_dat)
```

<br>

#### Plot 1

```{r LO3-EX1-plot1}
#percent of degrees conferred by year
ggplot(ba_dat, aes(year, perc)) + 
  geom_line(size=1) + 
  #make chart for each field 
  facet_wrap(~field) + 
  xlab("Year") + 
  ylab("Proportion of degrees")
```

<br>

#### Plot 2

```{r LO3-EX1-plot2}
#percent of degrees conferred by year
ggplot(ba_dat, aes(year, perc)) + 
   #colored area may make comparison between field easier  
  geom_area(alpha=0.5, fill="red") +
  #line on areaa (may want a different color )
  geom_line(color="red", size=1) + 
  #make chart for each field 
  facet_wrap(~field) + 
  xlab("Year") + 
  ylab("Proportion of degrees")
```


<br>

#### Plot 3

```{r LO3-EX1-plot3}
#percent of degrees conferred by year
ggplot(ba_dat
       , aes(
           year
         , perc
         #sample plot, differnet line color by field 
         , color=field 
         )
       ) + 
  geom_line(size=1) + 
  xlab("Year") + 
  ylab("Proportion of degrees") + 
  #change title on legend
  labs(color="Field") 
```

<br>

### Exercise 2

The following exercises use the `dow_jones-industrial` data set. 


```{r LO3-EX2-data, message=FALSE}
# Restrict data to useful range
djia_date_range <- dow_jones_industrial %>%
  filter(date >= ymd("2008/12/31") & date <= ymd("2010/01/10"))

#take a look at the data to see variable names and types
head(djia_date_range)
```

<br>

#### Plot 1


```{r LO3-EX2-plot1}
#closing price of DJIA by date 
ggplot(djia_date_range, aes(date, close)) + 
  geom_line(size = 1, color="purple") + 
  geom_smooth(
      method = "loess"
    , formula = y ~ x
    #line color
    , color = "green" 
    #se shading color 
    , fill = "red" 
    ) +
  xlab(NULL) + 
  ylab("Dow Jones Industrial Average")
```


<br>

#### Plot 2


```{r LO3-EX2-plot2}
#closing price of DJIA by date 
ggplot(djia_date_range, aes(date, close)) + 
  geom_line(size=1) + 
  geom_smooth(
      method="loess"
    , formula= y ~ x
    #wiggliness for loess curve  
    , span=0.3 
    #don't show se 
    , se=FALSE 
    ) + 
  xlab(NULL) + 
  ylab("Dow Jones Industrial Average")
```

<br>

#### Plot 3

```{r LO3-EX2-plot3}
ggplot(djia_date_range, aes(date, close)) + 
  geom_line(size=1) + 
  geom_smooth(
      method="lm"
      #splines makes the line more curvy and less 'chopped up' 
      #splines use piecewise polynomials
      #lowess is 'local regression'
    , formula= y ~ splines::ns(x, 6)
    , span=0.3
    , se = FALSE
    ) + 
  xlab(NULL) + 
  ylab("Dow Jones Industrial Average")
```


<br>

### Exercise 3

The following exercises use the `cdc` dataset.  

```{r}
#take a look at the data to see variable names and types
head(cdc)
```



#### Plot 1


**Using `cdc` and `geom_bar()**  
```{r LO3-EX3-plot1A}
ggplot(cdc, aes(genhlth)) + 
  #count genhlth 
  geom_bar()
```



**Using  `genhlth_count` and `geom_bar(stat="identity")`**

```{r LO3-EX3-plot1B}
#count genhlth and put into new data set 
genhlth_count <- cdc %>%
  count(genhlth)

#take a look at the data to see variable names and types
genhlth_count

ggplot(genhlth_count, aes(genhlth, n)) + 
  #use stat='identity' to plot actual values; rather than counting values 
  geom_bar(stat="identity") + 
  ylab("count")
```
**Using  `genhlth_count` and `geom_col()`**  
```{r LO3-EX3-plot1C}
ggplot(genhlth_count, aes(genhlth, n)) + 
  #geom_col will authomatically plot values 
  # where as geom_bar requires stat="identity" to do so 
  geom_col() + 
  ylab("count")
```

<br>

#### Plot 2  

```{r LO3-EX3-plot2}
ggplot(cdc, aes(genhlth, fill=as.factor(hlthplan))) + 
  geom_bar(position="dodge") + 
  #change title on legend
  labs(fill="Health Plan") 
```


<br>

#### Plot 3


```{r LO3-EX3-plot3}
ggplot(cdc
       , aes(
         #variable used to  calculate density 
           weight
          #color of area
         , fill = genhlth 
         #color of line
         , color = genhlth 
         )
       ) + 
  geom_density(alpha = 0.2) + 
  # create plots for each gender 
  facet_wrap(~gender)
```


<br>

#### Plot 4

The weight variable should have a lower limit of 50 and an upper limit of 300.  

**Fix x-axis limits**  
```{r LO3-EX3-plot4A, warning=FALSE}
ggplot(cdc, aes(weight, fill = gender, color = gender)) + 
  geom_density(alpha = 0.5) + 
  facet_wrap(~genhlth, scales = "fixed") + 
  xlim(50, 300)
```

**Filter cdc df to weight values [50, 300]**
```{r LO3-EX3-plot4B}
cdc %>%
  #filter data to only include values within the limts 
  filter(weight >= 50, weight <= 300) %>%
#pip directly into ggplot 
ggplot(aes(weight, fill = gender, color=gender)) + 
  geom_density(alpha=0.5) + 
  facet_wrap(~genhlth, scales = "fixed")
```


<br>

#### Plot 5

```{r LO3-EX3-plot5}
ggplot(cdc, aes(height, gender,  fill = gender)) + 
  geom_boxplot(alpha = 0.4) + 
  facet_wrap(~genhlth, nrow = 1) + 
  coord_flip()
```


<br>

#### Plot 6

```{r LO3-EX3-plot6}
ggplot(cdc, 
       aes(
           height
         , weight
         #put color here, inherited to geom_point and geom_smooth as well 
         , color = gender
         )
       ) + 
  geom_point(alpha = 0.2) + 
  geom_smooth(
      method = "lm"
    , formula = y~x
    , se = FALSE
    #extend line to cover whole plot 
    , fullrange = TRUE 
    )
```


