---
title: "L03 ggplot II"
subtitle: "Data Visualization (STAT 302)"
author: "Martha Eichlersmith"
date: 2020-04-19
output:
  html_document:
    toc: true
    toc_float: true
---

## Overview

The goal of this lab is to continue the process of unlocking the power of `ggplot2` through constructing and experimenting with a few basic plots.

## Datasets

We'll be using data from the `BA_degrees.rda` and `dow_jones_industrial.rda` datasets which are already in the `/data` subdirectory in our **data_vis_labs** project. Below is a description of the variables contained in each dataset.

`BA_degrees.rda`

* `field` - field of study
* `year_str` - academic year (e.g. 1970-71)
* `year` - closing year of academic year
* `count` - number of degrees conferred within a field for the year
* `perc` - field's percentage of degrees conferred for the year

`dow_jones_industrial.rda`

* `date` - date
* `open` - Dow Jones Industrial Average at open
* `high` - Day's high for the Dow Jones Industrial Average
* `low` - Day's low for the Dow Jones Industrial Average
* `close` - Dow Jones Industrial Average at close
* `volume` - number of trades for the day

We'll also be using a subset of the BRFSS (Behavioral Risk Factor Surveillance System) survey collected annually by the Centers for Disease Control and Prevention (CDC). The data can be found in the provided `cdc.txt` file --- place this file in your `/data` subdirectory. The dataset contains 20,000 complete observations/records of 9 variables/fields, described below.

* `genhlth` - How would you rate your general health? (excellent, very good, good, fair, poor)
* `exerany` - Have you exercised in the past month? (`1 = yes`, `0 = no`) 
* `hlthplan` - Do you have some form of health coverage? (`1 = yes`, `0 = no`) 
* `smoke100` - Have you smoked at least 100 cigarettes in your life time? (`1 = yes`, `0 = no`) 
* `height` -  height in inches
* `weight` - weight in pounds
* `wtdesire` - weight desired in pounds
* `age` - in years
* `gender` - `m` for males and `f` for females

<br>

```{r LO3-load-packages, warning=FALSE, message=FALSE}
# Load package(s)
library(tidyverse)
library(lubridate) #for ymd() function
```

## Exercises

Complete the following exercises.

<br>

### Exercise 1

Using `BA_degrees` dataset, recreate the following graphics as precisely as possible.

```{r LO3-EX1-data}
load(file="data/BA_degrees.rda")
# Wrangling for plotting
ba_dat <- BA_degrees %>%
  # mean % per field
  group_by(field) %>%
  mutate(mean_perc = mean(perc)) %>%
  # Only fields with mean >= 5%
  filter(mean_perc >= 0.05) %>%
  # Organizing for plotting
  arrange(desc(mean_perc), year) %>%
  ungroup() %>%
  mutate(field = fct_inorder(field))

glimpse(ba_dat)
```

<br>

##### Plot 1

```{r LO3-EX1-plot1}
ggplot(ba_dat, aes(year, perc))+
  geom_line(size=1)+
  facet_wrap(~field)+
  xlab("Year")+
  ylab("Proportion of degrees")
```



<br>

##### Plot 2

*Hints:*

- Transparency is 0.5
- Color used is`"red"`

```{r LO3-EX1-plot2}
ggplot(ba_dat, aes(year, perc))+
  geom_area(alpha=0.5, fill="red")+
  geom_line(color="red", size=1)+
  facet_wrap(~field)+
  xlab("Year")+
  ylab("Proportion of degrees")
```


<br>

##### Plot 3

```{r LO3-EX1-plot3}
ggplot(ba_dat, aes(year, perc, color=field))+
  geom_line(size=1)+
  xlab("Year")+
  ylab("Proportion of degrees")+
  labs(color="Field") #change title on legend
```

<br>

### Exercise 2

Using `dow_jones_industrial` dataset, recreate the following graphics as precisely as possible.  
*Hint:* Used `close`.

```{r LO3-EX2-data, message=FALSE}
load(file="data/dow_jones_industrial.rda")

# Restrict data to useful range
djia_date_range <- dow_jones_industrial %>%
  filter(date >= ymd("2008/12/31") & date <= ymd("2010/01/10"))
glimpse(djia_date_range)
```

<br>

##### Plot 1

*Hints:*

- Colors used `"red"`, `"purple"`, & `"green"`  
```{r LO3-EX2-plot1}
ggplot(djia_date_range, aes(date, close))+
  geom_line(size=1, color="purple")+
  geom_smooth(
      method="loess"
    , formula= y ~ x
    , color="green"
    , fill="red"
    )+
  xlab(NULL)+
  ylab("Dow Jones Industrial Average")
```


<br>

##### Plot 2

*Hints:*

- Wiggliness for `loess` is 0.3

```{r LO3-EX2-plot2}
ggplot(djia_date_range, aes(date, close))+
  geom_line(size=1)+
  geom_smooth(
      method="loess"
    , formula= y ~ x
    , span=0.3
    , se=FALSE
    )+
  xlab(NULL)+
  ylab("Dow Jones Industrial Average")
```

<br>

##### Plot 3

*Hints:*

- `y ~ ns(x, 6)` will need `splines` package (`"lm"` will work) 

```{r LO3-EX2-plot3}
ggplot(djia_date_range, aes(date, close))+
  geom_line(size=1)+
  geom_smooth(
      method="lm"
    , formula= y ~ splines::ns(x, 6)
    , span=0.3
    , se=FALSE
    )+
  xlab(NULL)+
  ylab("Dow Jones Industrial Average")
```


<br>

### Exercise 3

Using `cdc` dataset, recreate the following graphics as precisely as possible.

```{r LO3-EX3-data, warning=FALSE, message=FALSE}
# Read in the cdc dataset
cdc <- read_delim(file = "data/cdc.txt", delim = "|") %>%
  mutate(genhlth = factor(genhlth,
    levels = c("excellent", "very good", "good", "fair", "poor")
  ))
```

<br>

##### Plot 1

Construct this plot in two ways. Once using `cdc` and once using the `genhlth_count`.

**Using `cdc`**  
```{r LO3-EX3-plot1A}
ggplot(cdc, aes(genhlth))+
  geom_bar()
```



**Using  `genhlth_count`**

```{r LO3-EX3-plot1B}
genhlth_count <- cdc %>%
  count(genhlth)

glimpse(genhlth_count)

ggplot(genhlth_count, aes(genhlth, n))+
  geom_bar(stat="identity")+
  ylab("count")
```


<br>

##### Plot 2  

```{r LO3-EX3-plot2}
ggplot(cdc, aes(genhlth, fill=as.factor(hlthplan)))+
  geom_bar(position="dodge")+
  labs(fill="Health Plan") #change title on legend
```


<br>

##### Plot 3

*Hints:*

- Transparency is 0.2  

```{r LO3-EX3-plot3}
ggplot(cdc, aes(weight, fill=genhlth, color=genhlth))+
  geom_density(alpha=0.2)+
  facet_wrap(~gender)

```


<br>

##### Plot 4

*Hints:*

- Transparency is 0.5
- Horizontal axis should have lower limit of 50 and upper limit of 300  

**Fix x-axis limits**  
```{r LO3-EX3-plot4A, warning=FALSE}
ggplot(cdc, aes(weight, fill=gender, color=gender))+
  geom_density(alpha=0.5)+
  facet_wrap(~genhlth, scales="fixed")+
  xlim(50, 300)
```

**Filter cdc df to weight values [50, 300]**
```{r LO3-EX3-plot4B}
cdc %>%
  filter(weight >= 50, weight <= 300) %>%
ggplot(aes(weight, fill=gender, color=gender))+
  geom_density(alpha=0.5)+
  facet_wrap(~genhlth, scales="fixed")
```


<br>

##### Plot 5

*Hints:*

- Transparency is 0.4  

```{r LO3-EX3-plot5}
ggplot(cdc, aes(height, gender,  fill=gender))+
  geom_boxplot(alpha=0.4)+
  facet_wrap(~genhlth, nrow=1)+
  coord_flip()
```


<br>

##### Plot 6

*Hints:*

- Transparency is 0.2  

```{r LO3-EX3-plot6}
ggplot(cdc, aes(height, weight, color=gender))+
  geom_point(alpha=0.2)+
  geom_smooth(method="lm", formula= y~x, se=FALSE, fullrange=TRUE)
```


